---
title: "ggplot2: Secondary Y axis"
output: html_notebook
---

Use parameter `sec.axis` of `scale_y_continuous()`

# Simple example
```{r}
sapply(c("pipeR", "ggplot2", "readr", "lubridate"), require, character.only = TRUE)
```
Temperature and precipitation in Naha city, Okinawa, Japan (2015)

* Obtained from Japan meteorological agency
```{r}
d <- read_csv("data/ggplot2-two-axis/naha.csv", skip = 5, col_names = c("Date", "Temperature", "Precipitation"), col_types = cols(col_date(format = "%Y/%m"), "d", "_", "_", "d", "_", "_", "_"), locale = locale(encoding = "SJIS"))
head(d)
```
Precipitation
```{r}
d %>>% ggplot() + 
  geom_bar(mapping = aes(x = Date, y = Precipitation), stat = "identity") + 
  ylab("Precipitation (mm)")
```
Temperature
```{r}
d %>>% ggplot() + 
  geom_point(mapping = aes(x = Date, y = Temperature)) + 
  geom_line(mapping = aes(x = Date, y = Temperature)) + 
  ylab(expression("Temperature ("~degree~"C)"))
```
Range of data
```{r}
summary(d)
```
To set Temperature 0 -- 30 and Precipitation 0 -- 400, scale Precipitation by multiplying `30 / 400` to fit range of Temperature

```{r}
d %>>% ggplot() + 
  geom_bar(mapping = aes(x = Date, y = Precipitation * 30 / 400), stat = "identity") + 
  geom_point(mapping = aes(x = Date, y = Temperature)) + 
  geom_line(mapping = aes(x = Date, y = Temperature)) + 
  scale_y_continuous(name = expression("Temperature ("~degree~"C)"), limits = c(0, 30))
```

Scale first Y axis by multiplying `400 / 300` to create secondary Y axis for Precipitation

    * `scale_y_continuous(sec.axis = sec_axis(~ . * 400 / 30)) `

```{r}
d %>>% ggplot() + 
  geom_bar(mapping = aes(x = Date, y = Precipitation * 30 / 400), stat = "identity") + 
  geom_point(mapping = aes(x = Date, y = Temperature)) + 
  geom_line(mapping = aes(x = Date, y = Temperature)) + 
  scale_y_continuous(
    name = expression("Temperature ("~degree~"C)"),
    sec.axis = sec_axis(~ . * 400 / 30 , name = "Precipitation (mm)"), 
    limits = c(0, 30)) 
```

Final plot
```{r}
d %>>% ggplot() + 
  geom_bar(mapping = aes(x = Date, y = Precipitation * 30 / 400), stat = "identity", colour = gray(0.5), fill = gray(0.5)) + 
  geom_line(mapping = aes(x = Date, y = Temperature)) + 
  geom_point(mapping = aes(x = Date, y = Temperature), size = 3, shape = 21, fill = "white") + 
  scale_x_date(name = "Month", breaks = seq.Date(as.Date("2015-01-01"), as.Date("2015-12-31"), by = "1 month"), labels = function(date){return(month(date, label = TRUE))}) + 
  scale_y_continuous(
    name = expression("Temperature ("~degree~"C)"), 
    sec.axis = sec_axis(~ . * 400 / 30 , name = "Precipitation (mm)"), 
    limits = c(0, 30)) + 
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )
```

# Complex example: data contains negative values
Temperature and precipitation in Kushiro city, Hokkaido, Japan (2015)

* Obtained from Japan meteorological agency

```{r}
d2 <- read_csv("data/ggplot2-two-axis/kushiro.csv", skip = 5, col_names = c("Date", "Temperature", "Precipitation"), col_types = cols(col_date(format = "%Y/%m"), "d", "_", "_", "d", "_", "_", "_"), locale = locale(encoding = "SJIS"))
head(d2)
```
Precipitation
```{r}
d2 %>>% ggplot() + 
  geom_bar(mapping = aes(x = Date, y = Precipitation), stat = "identity") + 
  ylab("Precipitation (mm)")
```
Temperature
```{r}
d2 %>>% ggplot() + 
  geom_point(mapping = aes(x = Date, y = Temperature)) + 
  geom_line(mapping = aes(x = Date, y = Temperature)) + 
  ylab(expression("Temperature ("~degree~"C)"))
```

```{r}
summary(d2)
```

To set Temperature -5 -- 20 and Precipitation 0 -- 250: 
* Scale Precipitation by multiplying `1/10` to fit range of Temperature, after that, scale Precipitation by adding `-5`
* Scale first Y axis by adding `+5`, after that, scale Precipitation by multiplying `10` to create second Y axis for Precipitation

```{r}
d2 %>>% ggplot() + 
  geom_bar(mapping = aes(x = Date, y = Precipitation / 10 - 5), stat = "identity", colour = gray(0.5), fill = gray(0.5)) + 
  geom_line(mapping = aes(x = Date, y = Temperature)) + 
  geom_point(mapping = aes(x = Date, y = Temperature), size = 3, shape = 21, fill = "white") + 
  scale_x_date(name = "Month", breaks = seq.Date(as.Date("2015-01-01"), as.Date("2015-12-31"), by = "1 month"), labels = function(date){return(month(date, label = TRUE))}) + 
  scale_y_continuous(
    name = expression("Temperature ("~degree~"C)"), 
    sec.axis = sec_axis(~ (. + 5) * 10 , name = "Precipitation (mm)"), 
    limits = c(-5, 20)) + 
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )
```

`geom_bar()` start from `0`, use `geom_segment()`

```{r}
d2 %>>% ggplot() + 
  geom_segment(mapping = aes(x = Date, y = Precipitation / 10 - 5, xend = Date, yend = -5), size = 11, lineend = "butt", colour = gray(0.5)) +
  geom_line(mapping = aes(x = Date, y = Temperature)) +
  geom_point(mapping = aes(x = Date, y = Temperature), size = 3, shape = 21, fill = "white") +
  scale_x_date(name = "Month", breaks = seq.Date(as.Date("2015-01-01"), as.Date("2015-12-31"), by = "1 month"), labels = function(date){return(month(date, label = TRUE))}) +
  scale_y_continuous(
    name = expression("Temperature ("~degree~"C)"),
    sec.axis = sec_axis(~ (. + 5) * 10 , name = "Precipitation (mm)"),
    limits = c(-5, 20)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

```{r}
devtools::session_info()
```